### 2014-07-25 sections/abstract.tex

% !TEX root = ../paper.tex

\begin{abstract}
Software code review is an established method regarded as a best practice in Software Engineering to achieve a sufficient quality in software project.
Its main intention is to early identify defects before integrating changes.
Nowadays, Modern Code Review (MCR)\cite{Bacchelli2013a}, less formal and lightweight code reviews have received much attention and put into software development regularly in both industrial and OSS projects.
The main benefit of MCR is that in-person meetings are not required, unlike formal inspections \cite{Fagan:1976:DCI:1661010.1661012}.
Reviewers can find defects and discuss by adding comments through a code review tool or a mailing list.
Developers will then fix their changes following those comments.

% the freeform style of modern code review
Even though Modern Code Review has less process overhead, it can be less effective if not done properly.
\TODO{explain about our methods}

% some comments are not technically contributing

\begin{IEEEkeywords}
Modern Code Review, Software Quality, Text Mining
\end{IEEEkeywords}
\end{abstract}