### 2014-07-27 sections/introduction.tex
updated more of the introduction...

% !TEX root = ../paper.tex

\section{Introduction}
Software code review is an established cost-effective software quality management practice.
Its main intent is to  identify  defects early and encourage quality development practices such as adherence to coding standards and improving code readability.
Nowadays less formal and lightweight code reviews such as Modern Code Review (MCR)\cite{Bacchelli2013a} have received much attention and are commonly used within both industrial and OSS projects.
One reason for this is that with MCR, unlike formal inspections \cite{Fagan:1976:DCI:1661010.1661012}, in-person meetings are not required.
Reviewers can find defects and hold discussions on them by adding comments through a code review tool or a mailing list.
Developers then propose changes in response to these comments which in turn are reviewed for approval.  Fig. \ref{fig:process} illustrates a typical MCR process.

Because performance of code review has a clear relationship to the quality of software project, many studies have investigated the influencing factors in MCR\cite{Baysal2001,Mcintosh,Beller,Hamasaki2013}.
However, little research exists investigating how the discussion for a proposed change impacts software quality.
Given that most proposed changes are triggered from reviewer comments \cite{Beller}, 
it is possible that reviewers comments can either positively contribute to the proposed changes (i.e. ``\emph{useful}''), or be a discussion which is out of scope (i.e. ``\emph{useless}'' relative to the proposed changes\footnote{Although we call these comments \emph{useless}, it might be useful---or sometimes even required---in some other context, such as to facilitate communication.}).

Since MCR does not require strict guidelines, checklists, or formal processes, some people tend to review and just look for trivial issues (e.g. coding styles, conventions, and process issues), leaving deeper issues undiscussed \cite{Mcintosh}.
This is a concern because lack of useful comments can degrade the cost-effectiveness and perceived utility of MCR.

Current research has focused mostly on the relation between number of comments and quality.
To the best of our knowledge, the connection between degree of useful and useless comments and quality has not been studied.
One reason for this may be that to assess the impact of discussion in MCR, a manual classification is required as in the study of Beller et. al. \cite{Beller}.
MCR can produce a massive amount of changes and comments\cite{Balachandran2013,Thongtanunam2014}, and it is painstaking and time consuming to classify the comments in order to perform appropriate quantitative analysis.
Moreover, unlike a checklist in used formal inspection, the comments in MCR are unstructured natural text, open to subjective judgment, tacit understanding, and possible mis-interpretation.  Fig. \ref{fig:example} show typical MCR comments illustrating these issues.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35, trim= 100 110 50 80, clip=true]{review_process}
\caption{An simplified version of MCR Process based on Gerrit system}
\label{fig:process}
\end{figure}

\dan{Much of the following material should be placed in a separate Motivation section} 
In this paper we present a text mining approach to assist in classifying the usefulness of comments.
The proposed approach is meant to enable classifying the large amounts of MRC comments, in order to conduct comment and code quality studies, such as those discussed earlier, as manual classification can quickly become an untenable task.

Our modest OSS case study has a over 17,000 potentially useful comments.
At an optimistic 5 seconds peer comment classification, we would expect about 24 person-hours of effort.
Furthermore, manual classification is subjective and error-prone.

By introducing some automation support based on objective empirical criteria,
the method we propose here aims to both reduce effort and increase confidence in classification to a degree that is practical for addressing very large MRC data sets.
Because usefulness of a comment is not precisely defined and is a subjective qualification, we cannot, and do not intend to fully automate or replace human driven classification.

We define a \emph{useful} comment as one that \emph{contributes to improving the proposed changes} and a \emph{useless} comment as one that does not.
Usefulness is not dichotomous and hence we must accept a third qualification, \emph{unclear}, to allow for the case where a comment does not directly make a positive contribution but is not clearly out of scope or tangential.

In our approach, we analyze the similarity between a commit message of the proposed changes and the respective comments that follow.
Our key idea is that useful comments are likely to have a strong semantic similarity to the proposed changes, while useless comments are more likely strongly dissimilar.
To determine these semantic relationships, we use the Vector Space Model (VSM) and calculate similarity (using cosine similarity) and dissimilarity (using euclidean distance) between the comment and commit message vectors. 
We create our predictive model by estimating pairs of similarity and dissimilarity values that best discriminate between useful and useless comments.
This is accomplished by optimizing for prediction performance measures such as Precision and Recall, through F-measure.

For our case study of the method, we used a review history from the Gerrit\footnote{\url{https://code.google.com/p/gerrit/}} system from Qt\footnote{\url{http://qt-project.org/}}, an open source project for a cross-platform application and UI framework supported by Digia corporation.
For training and validation of our approach, three experts manually classified the usefulness for 320 comments.
We also apply our predictive model to the full set of 17,000 \dan{what's the actual number?}
comments to obtain a preliminary estimate in the quality of reviews and address the following three research questions:

\noindent \textbf{RQ1:} Is semantic similarity a good indicator of MCR comment usefulness?\\
\noindent \textbf{RQ2:} Do code reviewers intensively discuss on the proposed changes?\\
\noindent \textbf{RQ3:} Is semantic similarity classification cost-efficient, assurable, and scalable?

\noindent In addition to addressing the research questions above, the main contributions of this paper are:
\dan{these are more "results" than contributions.... need to reword this!}
\begin{itemize}
\item A practical to apply approach to mine natural text of comments in code reviews and classify their usefulness based on VSM similarity.
\item The experimental results show that the approach can classify comments with an F$_1$ score of 0.681 and 0.693.
\item We found that 85\% of comments in Gerrit are automatically generated and not likely useful (but important for record keeping). For the remaining 15\%, about 43\% of them are automatically classified as useful, while about 34\% are classified as not useful. The remaining 23\% are undetermined and require manual classification. Thus the classification effort is reduced by about 77\%.
\item xx\% of the comments did not have unanimity (i.e. unclear), and of these yy\% were automatically determined to be useful or not useful. Follow up with dissenting assessors found mis-judgement occurred zz\% of the time whereby in retrospect the comment would have had unanimity. This indicates the method is useful in identifying subjective judgement errors. 
\item Of the 23\% not automatically classified (i.e. undetermined), qq\% were unclear. Of the (1-qq)\% with unanimity, followup with assessors found in retrospect they would change their assessment making the comment unclear. This indicates the method is useful in identifying chance unanimity errors or group think bias.
\end{itemize} 
%Software Inspection is basically composed of a three-step procedure: preparation, inspection meeting, and repair.