### 2014-07-29 paper.tex
add conclusion

\documentclass[conference]{IEEEtran}
\usepackage{microtype,mathtools,amsmath,multibib,amsfonts,multicol,array,multirow,place ins,cite,balance,makecell,algorithm,algorithmic,subfigure,paralist}
%\usepackage{flushend}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{url}
\usepackage[font=small,labelfont=bf]{caption}

\usepackage{xspace}
\usepackage{color}
\usepackage{ifthen}
\usepackage{fancybox}

\let\labelindent\relax
\usepackage{enumitem}
\newlist{ResearchQuestions}{description}{2}
\setlist[ResearchQuestions]{leftmargin=3em,style=nextline}

\graphicspath{{zu/}}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{Assessing MCR Discussion Usefulness using  Semantic Similarity}
%\title{Using Semantic Similarity to Assist Assessing MCR Comment Usefulness }

\author{
\IEEEauthorblockN{Thai Pangsakulyanont\IEEEauthorrefmark{1}, Patanamon Thongtanunam\IEEEauthorrefmark{2},  Daniel Port\IEEEauthorrefmark{3}, Hajimu Iida\IEEEauthorrefmark{2}}
\\
\IEEEauthorblockA{
\IEEEauthorrefmark{1} Kasetsart University, Thailand\\
b5410545036@ku.ac.th
}

\IEEEauthorblockA{
\IEEEauthorrefmark{2} Nara Institute of Science and Technology, Japan\\
patanamon-t@is.naist.jp, iida@itc.naist.jp\\
}

\IEEEauthorblockA{
\IEEEauthorrefmark{3} University of Hawaii at Manoa, USA\\
dport@hawaii.edu \\
}
}

\maketitle
\newboolean{showcomments}
\setboolean{showcomments}{true} % toggle to show or hide comments
\ifthenelse{\boolean{showcomments}}
{\newcommand{\nbnote}[2]{
  % \fbox{\bfseries\sffamily\scriptsize#1}
  \fcolorbox{blue}{yellow}{\bfseries\sffamily\scriptsize#1}
  {\sf\small\textit{#2}}
  % \marginpar{\fbox{\bfseries\sffamily#1}}
 }
}
{\newcommand{\nbnote}[2]{}
 \newcommand{\version}{}
}
\newcommand\dan[1]{\nbnote{Dan sez}{\textcolor{red}{#1}}}
\newcommand\pick[1]{\nbnote{Pick sez}{\textcolor{magenta}{#1}}}
\newcommand\thai[1]{\nbnote{Thai sez}{\textcolor{blue}{#1}}}
\newcommand\TODO[1]{\textcolor{red}{\textbf{TODO} #1}}
\newcommand\True{\operatorname{True}}
\newcommand\False{\operatorname{False}}
\newcommand\Sim{\operatorname{sim}}
\newcommand\Dist{\operatorname{dist}}

\input{sections/abstract}
\input{sections/introduction}
\input{sections/background}
\input{sections/approach}
%\input{sections/methods}
\input{sections/results}
\input{sections/discussion}
\input{sections/conclusion}

\IEEEpeerreviewmaketitle

\bibliographystyle{IEEEtran}

\bibliography{references}

% that's all folks
\end{document}