### 2014-07-29 sections/conclusion.tex

% !TEX root = ../paper.tex

\section{Conclusion and Future Work}
In this paper, we propose a practical approach to classify usefulness discussion in MCR based on VSM similarity.
The results from our case study show that our approach can classify comments achieved better results than random model twice for both precision and recall.
Moreover, we found that 85\% of comments in Gerrit are automatically generated and not likely useful (but important for record keeping). For the remaining 15\%, about 43\% of them are automatically classified as useful, while about 34\% are classified as not useful. The remaining 23\% are undetermined and require manual classification. Thus the classification effort is reduced by about 77\%.

For our future work,.....