### 2014-07-30 sections/conclusion.tex
forgot to commit

% !TEX root = ../paper.tex

\section{Conclusions}